\bf Calendar-based graphics for visualizing people's daily schedules

### 2019-10-18 reviews/response.Rmd

We thank the reviewers for their helpful feedback which has improved the paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

# Associate editor {-}

> What if the data's collected over different time zones?

We added a short paragraph in Section 2.1 after "Note that for the vertical direction" to address the time zone problem.

"The R package, **lubridate**, is used to extract the components of time, such as days of the week and the number of days in a month, that create the layout. These time variables are converted to integers for the modular arithmetic. Note that for any date-time information is associated with time zone. If your data is collected over multiple time zones, you will need to convert them to the same time zone before conducting any temporal analysis."

@@

Done.

# Reviewer 1 {-}

> "Would it be possible to have a plot for each day in a calendar plot with thick lines separating the months?  That kind of plot is common for calendar heatmaps."  I think that is important.  Perhaps they can build it in.  If not, why not?

The white space between months already emphasises the separations between months. The lines separating the months are also thicker than the ones separating the days in the calendar plot. These sit in the background to make the data plot more visible, which follows the Tufte's rules.

@@

> p10 Figure 6  Could the vertical scale be given in the caption?

We added the range of vertical scale in the caption.

> p14 and 15-16 Figure 9 is better than Figure 8.  Are both necessary?

We kept both figures in, to show what users can do with the calendar plots, for example both overlaying and faceting.

@@

> p14 "Almost any kind of plot can be shown in a calendar pane."  This is excellent.  An impressive example would be useful.  Figure 10 is disappointing.  It is difficult to read and the information it provides could be seen in earlier figures.

We have rephrased this sentence to "Many types of plot can be shown in a calendar pane, by taking advantage of the existing **ggplot2** plotting capabilities.". You have inspired us to try to make more exciting calendar plots, and they are possible but not useful. The lag scatterplots (now figure 8) is the sensible data plot for the pedestrian dataset, that fits into the Section 2.3.3. To make the information easier to read, we plot a subset from week 1 to 17 instead of the whole year before.

@@

> p17 Figure 10  Why not align by days of the week?

We have changed figure 10 (now figure 8) to align by days of the week instead of by days of the month. Since the best display for the pedestrian dataset is monthly layout, we wanted to showcase other possible layouts.

@@

> The case study is interesting.  Why are the data for household 2 different (fewer possible values)?

Data is of the same time lengths (and the same amount of data points) across four households. Different households have different lifestyles.

> p22 "with noticeable variations on Thursdays" based on only 26 data points is over-interpretation.

It's half-hourly data, meaning 26 Thursdays * 48 per day = 1248 data points. But we dropped the boxplot as in figure 13 including this sentence.

### 2019-10-18 calendar-vis.Rmd
addressed all the comments

The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), variations of its usage (Section \@ref(sec:variations)), and the full faceting extension equipped with formal labels and axes (Section \@ref(sec:facet-calendar)). An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.

@@

\noindent Figure \@ref(fig:month-diagram) illustrates this $(i, j)$ layout for a month where $k = 5$.

@@

@@

Almost any type of plot can be shown in a calendar pane. Most of the full range of plotting capabilities in **ggplot2** is available. An example is shown in Figure \@ref(fig:scatterplot): the panes contain lag scatterplots, constructed with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. It indicates strong autocorrelation on weekends, and weak autocorrelation on work days. The V-shape in the weekday graphs arises when the next hour sees a substantial increase or decrease in counts.

```{r scatterplot, fig.height = 12, out.width = "75%", fig.cap = "(ref:scatterplot-cap)"}
```

(ref:scatterplot-cap) Lag scatterplots, with local scaling, in the weekly calendar layout. Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days (blue) than work days (red).

>>>>

The remainder of the paper is organized as follows. Section \@ref(sec:algorithm) details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \@ref(sec:transformation)), the available options (Section \@ref(sec:opt)), variations of its usage (Section \@ref(sec:variations)), including the full faceting extension equipped with formal labels and axes (Section \@ref(sec:facet-calendar)). An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \@ref(sec:case). Section \@ref(sec:discussion) discusses the limitations of calendar displays and possible new directions.

@@

@@

The R package, **lubridate** [@R-lubridate], is used to extract the components of time, such as days of the week and the number of days in a month, that create the layout. These time variables are converted to integers for the modular arithmetic. Note that for any date-time information is associated with time zone. If your data is collected over multiple time zones, you will need to convert them to the same time zone before conducting any temporal analysis.

@@

Many types of plot can be shown in a calendar pane, by taking advantage of the existing **ggplot2** plotting capabilities. An example is shown in Figure \@ref(fig:scatterplot): the panes contain lag scatterplots for Flagstaff Station from Week 1 to 17 in 2016, constructed with the scaling for each day and aligning by days of the week, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. It indicates strong autocorrelation on weekends, and weak autocorrelation on work days. The V-shape in the weekday graphs arises when the next hour sees a substantial increase or decrease in counts.

```{r scatterplot, fig.height = 12, out.width = "70%", fig.cap = "(ref:scatterplot-cap)"}
```

(ref:scatterplot-cap) Examining lag 1 autocorrelation for Flagstaff Station from week 1 to 17, using lag scatterplots scaled by each day and aligned by days of the week. Each hour's count is plotted against the previous hour's count. The autocorrelation is stronger on non-work days (blue) than work days (red).

### 2019-10-18 calendar-vis.tex
addressed all the comments

The remainder of the paper is organized as follows. Section \ref{sec:algorithm} details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \ref{sec:transformation}), the available options (Section \ref{sec:opt}), variations of its usage (Section \ref{sec:variations}), and the full faceting extension equipped with formal labels and axes (Section \ref{sec:facet-calendar}). An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.

@@

The algorithm of transforming data for constructing a calendar plot uses modular arithmetic, similar to that used in the glyph map displays for spatio-temporal data \citep{Wickham2012glyph}. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to weeks of the month and days of the week respectively. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar. Months are of different lengths, ranging from 28 to 31 days. Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention. The notation for creating these cells is as follows:

@@

\noindent Figure \ref{fig:month-diagram} illustrates this \((i, j)\) layout for a month where \(k = 5\).

@@

@@

Almost any type of plot can be shown in a calendar pane. Most of the full range of plotting capabilities in \textbf{ggplot2} is available. An example is shown in Figure \ref{fig:scatterplot}: the panes contain lag scatterplots, constructed with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument. It indicates strong autocorrelation on weekends, and weak autocorrelation on work days. The V-shape in the weekday graphs arises when the next hour sees a substantial increase or decrease in counts.

@@

{\centering \includegraphics[width=0.75\linewidth]{figure/scatterplot-1}

@@

\caption{Lag scatterplots, with local scaling, in the weekly calendar layout. Each hour's count is plotted against the previous hour's count at Flagstaff Station, to demonstrate the autocorrelation at lag 1. The correlation between them is more consistent on non-work days (blue) than work days (red).}\label{fig:scatterplot}
\end{figure}

>>>>

The remainder of the paper is organized as follows. Section \ref{sec:algorithm} details the construction of the calendar layout in depth. It describes the algorithms of data transformation (Section \ref{sec:transformation}), the available options (Section \ref{sec:opt}), variations of its usage (Section \ref{sec:variations}), including the full faceting extension equipped with formal labels and axes (Section \ref{sec:facet-calendar}). An analysis of half-hourly household energy consumption, using the calendar display, is illustrated in a case study in Section \ref{sec:case}. Section \ref{sec:discussion} discusses the limitations of calendar displays and possible new directions.

@@

The algorithm of transforming data for constructing a calendar plot uses modular arithmetic, similar to that used in the glyph map displays for spatio-temporal data \citep{Wickham2012glyph}. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month conforms to a layout of 5 rows and 7 columns, where rows and columns refer to weeks of the month and days of the week respectively. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is deterministic given the year of the calendar. Months are of different lengths, ranging from 28 to 31 days. Some months could extend over six weeks, but for these months the last few days are wrapped up to the top row of the block for compactness, and because it is convention. The fifth row could be blank for February if the month starts on Monday. The notation for creating these cells is as follows:

@@

@@

The R package, \textbf{lubridate} \citep{R-lubridate}, is used to extract the components of time, such as days of the week and the number of days in a month, that create the layout. These time variables are converted to integers for the modular arithmetic. Note that for any date-time information is associated with time zone. If your data is collected over multiple time zones, you will need to convert them to the same time zone before conducting any temporal analysis.

@@

Many types of plot can be shown in a calendar pane, by taking advantage of the existing \textbf{ggplot2} plotting capabilities. An example is shown in Figure \ref{fig:scatterplot}: the panes contain lag scatterplots for Flagstaff Station from Week 1 to 17 in 2016, constructed with the scaling for each day and aligning by days of the week, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument. It indicates strong autocorrelation on weekends, and weak autocorrelation on work days. The V-shape in the weekday graphs arises when the next hour sees a substantial increase or decrease in counts.

@@

{\centering \includegraphics[width=0.7\linewidth]{figure/scatterplot-1}

@@

\caption{Examining lag 1 autocorrelation for Flagstaff Station from week 1 to 17, using lag scatterplots scaled by each day and aligned by days of the week. Each hour's count is plotted against the previous hour's count. The autocorrelation is stronger on non-work days (blue) than work days (red).}\label{fig:scatterplot}
\end{figure}
